---
title: "clustering"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(zoo)
library(tidyverse)
library(readxl)
library(tibble)
library(janitor)
library(reshape2)
library(ggplot2)
library(ggthemes)
library(plotly)
library(pheatmap)
library(cluster)
library(factoextra)
```

Reading in a matrix-ed version of the new cases, available on google drive.


```{r}
df = read.csv("new_cases_matrix.csv")
head(df)
```

This data is no NAs.

```{r}
sum(is.na(df))
```
```{r}
colSums(dplyr::select(df, -date))
```




## Clustering
 
Since Turkmenistan,  Kiribati, and Tonga have zero cases reported, we remove them from the dataset as the normalisation would yield infinity. 


```{r}
new_cases <- df %>% dplyr::select(-date, -Turkmenistan, -Kiribati, -Tonga)
```


```{r}
new_cases
```

```{r}
new_case_matrix_norm <- apply(new_cases, 2, function(x) x/sum(x))
new_case_matrix_norm_dist <- dist(t(new_case_matrix_norm), method = "manhattan")
```

```{r}
hclust_res <- hclust(new_case_matrix_norm_dist, method="ward.D2")
hclust_cluster <- cutree(hclust_res, k = 15) %>% as.factor %>% as.data.frame 
```

An interactive plot isn't particularly helpful here because we have SO many countries together. 

```{r}
df_covid <- reshape2::melt(new_case_matrix_norm)
colnames(df_covid) <- c("date", "country_name", "new_confirmed")
df_covid <- df_covid %>% mutate(date = as.Date(date))
df_covid$cluster <-  hclust_cluster[as.character(df_covid$country_name), 1]
g = ggplot() +
  geom_line(data = df_covid, aes(x = date, y = new_confirmed, color = country_name)) +
  facet_wrap(~cluster, ncol = 1) +  ## argument scale allow you to adjust whether to have the same y-axis or not
  theme_bw() + ylab("Normalised values") + labs(title = "COVID new cases")
plotly::ggplotly(g)
```


Let's try plotting the hclust as a dendrogram

```{r}
plot(hclust_res, cex=0.6)
```


```{r}
hclust_cluster <- cutree(hclust_res, k = 5) %>% as.factor %>% as.data.frame 
colnames(hclust_cluster) = "clust_res"
pheat <- pheatmap(t(new_case_matrix_norm), 
         cluster_cols = FALSE,
         main = "New cases",
         annotation_row = hclust_cluster,
         clustering_distance_rows = "manhattan",
         clustering_method = "ward.D2",
         
         cex=0.5)
```
This is virtually incomprehensible, so let's take a look at individual clusters ay


```{r}
CLUST_NUM = 15
g = ggplot() +
  geom_line(data = df_covid[df_covid$cluster == CLUST_NUM,], aes(x = date, y = new_confirmed, color = country_name)) +
  facet_wrap(~cluster, ncol = 1) +  ## argument scale allow you to adjust whether to have the same y-axis or not
  theme_bw() + ylab("Normalised values") + labs(title = "COVID new cases")
plotly::ggplotly(g)
```




# Functionalising it to test some parameters quickly

Overall visualisations
- Small versions of all curves if possible
- Select cluster number to view
- Summary of all clusters, how many countries in each, and a preview of which countries are in it


**We Will Also Add Smoothing Or Some Kind Of Standardisation**


```{r}
# Minmax scaling 
min_max_scale = function(series) {
  return ((series - min(series))/(max(series) - min(series)))
}
new_cases_scaled <- apply(new_cases, 2, min_max_scale)
```


And rerunning clustering to get our dendrogram!


```{r}
new_cases_scaled_dist <- dist(t(new_cases_scaled), method = "manhattan")
hclust_scaled_res <- hclust(new_cases_scaled_dist, method="ward.D2")
hclust_scaled_cluster <- cutree(hclust_res, k = 15) %>% as.factor %>% as.data.frame 
colnames(hclust_scaled_cluster) = "clust_num"
```

And here's our dendrogram!

```{r}
plot(hclust_scaled_res, cex=0.6)
```

```{r}
hclust_scaled_cluster %>% group_by(clust_num) %>% count()
```
So Cluster 8, 14, and 15 have the fewest countries in the cluster, while Cluster 4 has the most. We'll take a closer look to figure out why. 


```{r}
df_covid_scaled <- reshape2::melt(new_cases_scaled)
colnames(df_covid_scaled) <- c("date", "country_name", "new_confirmed_scaled")
df_covid_scaled$date <- as.Date(df$date)
df_covid_scaled$cluster <-  hclust_scaled_cluster[as.character(df_covid_scaled$country_name), 1]
```

```{r}
df_covid_scaled
```


```{r}
which(new_cases_scaled[1,] %in% c("Brunei"))
```
```{r}
data.frame(new_cases_scaled)
```




```{r}
CLUST_NUM = 3
g = ggplot() +
  geom_line(data = df_covid_scaled[df_covid_scaled$cluster == CLUST_NUM,], aes(x = date, y = new_confirmed_scaled, color = country_name)) +
  facet_wrap(~cluster, ncol = 1) +  ## argument scale allow you to adjust whether to have the same y-axis or not
  theme_bw() + ylab("Normalised values") + labs(title = "COVID new cases (scaled)")
plotly::ggplotly(g)
```


```{r}
a <- df %>% dplyr::select( -Turkmenistan, -Kiribati, -Tonga) 
a$date = as.Date(a$date)
```


```{r}
#Gap Statistic Method

gap_stat <- clusGap(new_cases_scaled, FUN = hcut, nstart = 25, K.max = 20, B = 50)
fviz_gap_stat(gap_stat)
```

```{r}
#Elbow Method

fviz_nbclust(new_cases_scaled, FUN = hcut, method = "wss")
```

```{r}
#Silhouette Method

fviz_nbclust(new_cases_scaled, FUN = hcut, method = "silhouette")
```

```{r}

```