---
title: "Lag_time"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(plotly)
library(zoo)

```

# Lag Time Investigation

Objectives:
1. Determine a suitable lag time measure between government response and new covid cases.
2. Identify average lag time for different countries and across time.


## 1. Reading in the Cleaned Dataset

```{r}
df = read.csv("epi_gov_data_cleaned.csv")

head(df)


```

Because cases can vary from 0 to like 500k, while stringency index is only ranged from 0 to 100, we'll standardise/normalise the data so they can be plotted on the same graph. 

```{r}
min_max_scale = function(series) {
  return ((series - min(series))/(max(series) - min(series)))
}
```



```{r}
# Minmax scaling
df$new_confirmed_minmax = df$new_confirmed - min(df$new_confirmed)


df = df %>%
  group_by(country_name) %>%
  mutate(
    new_confirmed_minmax = min_max_scale(new_confirmed),
    stringency_minmax = min_max_scale(stringency_index)
    
  )

```



## 2. Intial Correlation and Lag musings

Referencing: https://online.stat.psu.edu/stat510/lesson/8/8.2
- Cross Correlation Functions (CCF) is helpful for identifying lags of the x-var that might be useful in predicting y
- Tests many combinations of lag time units (h=0, +-1, +- 2, etc)
- I believe this assumes x is the sole predictor of y, which is also fine because for the most part, it makes sense for government response to be determined by the number of covid cases. 


Plotting cases against stringency as an initial trial. 
```{r}

# Plotting our cases against government responses first
df_sample = df[df$country_name %in% c("Australia", "United States of America", "Singapore", "Mexico", "United Kingdom", "China"),]

plt = ggplot() + 
  geom_line(data=df_sample, aes(x=date, y=(stringency_index/100), group=country_name,
                         color=country_name), color='red') + 
  geom_line(data=df_sample, aes(x=date, y=new_confirmed_minmax, group=country_name,
                         color=country_name), color="blue") +
  facet_wrap(~country_name, scale="free_y") + 
  ylab("Number of cases vs stringency (min-max scaled)") + 
  labs(color = "Country") + 
  ggtitle("New cases and government stringency by country")

plotly::ggplotly(plt)

```

We do find some level of association on a macro level, some more than others. E.g. Australia and Singapore seem to be more variable with their adjustments, lowering when cases are low and raising them when cases are high.

USA and Mexico on the other hand, seem to have maintained high stringency all throughout. 


First doing a study on Australia:

```{r}
# Setting up time series

aus_df = df[df$country_name == "Australia",]

cases = ts(aus_df$new_confirmed)
stringency = ts(aus_df$stringency_index)

ccf(cases, stringency)
```

```{r}
ccf(cases, stringency,
                 main="Cases vs Stringency ACF Lag", xlab="Lag (days)")

```


```{r}

library(forecast)
ggCcf(cases, stringency, main="AAH")


```





Running CCF on Australia gives us a very wide range of possible lag times. It peaks somewhere around the h= -10 mark, suggesting about a week of lag between cases and stringency. 

Let's run it again on a min max scaled cases

```{r}
ccf(ts(aus_df$new_confirmed_minmax), ts(aus_df$stringency_index))

```


And with log transformation so the big cases are less spikey. Because we have zero cases in the dataset, we'll do a log (cases + 1), which has minimal impact on the scaling.


```{r}
ccf(log(cases+1), stringency)
```

Ahk Lag aint too clear but it's also likely because stringency has a lower propensity to change in the short term-- i.e. once you increase/decrease stringency, it will likely remain that way for the next few days.

```{r}
ccf_values = ccf(cases, stringency)
ccf_values
```

If we check out the ccf values themselves, which contains the correlation for each time lag, we can see that a 12 day lag has the highest correlation but they're very similar around the 7 - 16 day lag. 

Meaning that on average over the 2020 year, Australia's government takes about 1-2 weeks to respond to covid cases.

Key word being over the 2020 year. 

### Prelim lag over months

Our hypothesis is that Government's response to covid have changed over the course of the year.

So now we'll try segmenting the data by month/quarter and calculate lag on that

```{r}
ccf_values = ccf(
  ts(aus_df[aus_df$date < "2020-04-01",]$new_confirmed),
  
  ts(aus_df[aus_df$date < "2020-04-01",]$stringency_index)
  )
ccf_values


```

```{r}
ccf_values = ccf(
  ts(aus_df[(aus_df$date >= "2020-03-01")&(aus_df$date < "2020-04-01"),]$new_confirmed),
  
  ts(aus_df[(aus_df$date >= "2020-03-01")&(aus_df$date < "2020-04-01"),]$stringency_index)
  )
ccf_values
```




```{r}
ccf_values = ccf(
  ts(aus_df[(aus_df$date >= "2020-04-01")&(aus_df$date < "2020-07-01"),]$new_confirmed),
  
  ts(aus_df[(aus_df$date >= "2020-04-01")&(aus_df$date < "2020-07-01"),]$stringency_index)
  )
ccf_values

```


```{r}
ccf_values = ccf(
  ts(aus_df[(aus_df$date >= "2020-07-01")&
              (aus_df$date < "2020-10-01"),]$new_confirmed),
  
  ts(aus_df[(aus_df$date >= "2020-07-01")&
              (aus_df$date < "2020-10-01"),]$stringency_index)
  )
ccf_values
```


```{r}
ccf_values = ccf(
  ts(aus_df[(aus_df$date >= "2020-10-01")&
              (aus_df$date < "2021-01-01"),]$new_confirmed),
  
  ts(aus_df[(aus_df$date >= "2020-10-01")&
              (aus_df$date < "2021-01-01"),]$stringency_index)
  )
ccf_values
```


HMMM ok so over the course of 2020 quarters, we see this relationship:
- First quarter (Jan - Mar): 0 day lag with 80% correlation
- Second quarter (Apr - Jun): 8 - 9 days, with 40% correlation
- Third quarter (Jul - Sep): 12 - 13 day reverse lag with -68% correlation. This means govt stringency actually came before cases. Negative correlation means more stringent --> cases went down 2 weeks later. 
- Fourth quarter (Oct - Dec): 1 day lag. 


Seems interesting, but we also note that covid cases basically went on a downspike during the third period, which may explain the reversal in lag. 
Other issues with correlation coefficients (simple)
- Not robust to outliers (which we have a lot of in covid cases)
- Assumes homoscedasticity = variance homogenous (which it is not)

But still, with more window sizes and bins we might be able to build a better picture of lag time. 

```{r}
ccf_values = ccf(
  ts(aus_df[(aus_df$date >= "2021-01-01"),]$new_confirmed),
  
  ts(aus_df[(aus_df$date >= "2021-01-01"),]$stringency_index)
  )
ccf_values
```


## 3. ROLLing correlation

BECAUSE static correlation is not appropriate. 

We use `zoo::rollapply`


```{r}

roll_corr_monthly = rollapply(aus_df[c("new_confirmed", "stringency_index")],
          width=30, 
          function(x) cor(x[,1], x[,2]), 
          by.column=FALSE)

```


```{r}
aus_df$date = as.Date(aus_df$date)
```


```{r}
date_x = aus_df$date[30:nrow(aus_df)]
```


```{r}
plot(roll_corr_monthly)
```


```{r}
plot(date_x, roll_corr_monthly, 'l')
```

```{r}
roll_corr_df = data.frame(list(date=date_x, roll_corr = roll_corr_monthly))
```


```{r}
ggplot() +
    geom_hline(yintercept=0) + 
  geom_point(data=roll_corr_df, aes(x=date_x, y=roll_corr_monthly),
    color="blue") +

  theme_minimal() + 
  ggtitle("Rolling correlation over past 30 days")
```

Ok so over the course of the month is probably not the best way to do this. Let's try 90 days and see if we can get a match

```{r}
plot_rolling_corr = function(ndays=30, df) {
  df$date = as.Date(df$date)
  roll_corr_calc = rollapply(df[c("new_confirmed", "stringency_index")],
          width=ndays, 
          function(x) cor(x[,1], x[,2]), 
          by.column=FALSE)
  date_ndays = df$date[ndays:nrow(df)]
  
  roll_corr_df = data.frame(list(date=date_ndays, roll_corr = roll_corr_calc))
  
  ggplot() +
  geom_line(data = roll_corr_df, aes(x=date_ndays, y=roll_corr_calc), color="blue") +
      geom_hline(yintercept=0, color='lightgrey') + 
  theme_minimal() + 
  ggtitle(paste("Rolling correlation over past", ndays, "days"))

}
```


```{r}
plot_rolling_corr(90, aus_df)
```


Next Steps:
- Rolling correlation over multiple time periods as interactive thing
- Interactive plot calculating lag time in days
^ Both stratified by country and time period


```{r}

us_df = df[df$country_name == "United States of America",]

ccf(ts(us_df$new_confirmed), ts(us_df$stringency_index)
)


```

```{r}
us_df = df[df$country_name == "France",]

ccf(ts(us_df$new_confirmed), ts(us_df$stringency_index)
)

```







Resources:
- https://www.listendata.com/2017/04/how-to-standardize-variable-in-regression.html
- https://towardsdatascience.com/four-ways-to-quantify-synchrony-between-time-series-data-b99136c4a9c9
- https://nwfsc-timeseries.github.io/atsa-labs/sec-tslab-decomposition-of-time-series.html
- https://www.statology.org/rolling-correlation-in-r/
- 






