---
title: "som"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(kohonen)
library(tidyverse)

set.seed(123)
```


```{r}
df = read.csv("new_cases_matrix.csv", row.names="date", header=TRUE)

```



## Clustering
 


We need rows to be the thing to group on (country) and columns to be features (in this case, date?)



Since Turkmenistan,  Kiribati, and Tonga have zero cases reported, we remove them from the dataset as the normalisation would yield infinity. 


```{r}
normalize <- function(x, na.rm = TRUE) {
    return((x- min(x)) /(max(x)-min(x)))
}

```



```{r}
new_cases <- data.frame(apply(
        df %>% dplyr::select(-Turkmenistan, -Kiribati, -Tonga)
        , 2, normalize))
```


```{r}
new_cases <- t(new_cases)
```




```{r}
som_model <- som(
        new_cases, 
        grid=somgrid(xdim = 4, ydim=4, 
                     topo="hexagonal"), 
        rlen=100, 
        alpha=c(0.05,0.01), 
        keep.data = TRUE,)
```




Plotting the training progress, which should decrease over time until it plateaus

```{r}
plot(som_model, type="changes")
```

Node counts indicate how many samples in the dataset are mapped to the same node. This should be relatively uniform-- otherwise may need a bigger map. Here, we see one node has 4x the number of countries as others


```{r}
plot(som_model, type="count")
```



Code referencing: https://clarkdatalabs.github.io/soms/SOM_Shakespeare_Part_1
and https://www.r-bloggers.com/2014/02/self-organising-maps-for-customer-segmentation-using-r/


```{r}
plot(som_model,
     type = "mapping",
     # bgcol = "lightgray",
     labels = rownames(new_cases))
```

Here now we only display the countries we're interested in seeing





```{r}



to_plot <- c("United.States.of.America", 
             "Italy", "Iran", "France", "Australia", "New.Zealand",
             "India", "Brazil", "China", "South.Korea", "Japan", "United.Kingdom", "Russia")

new_labs <- c()
X <- as.vector(rownames(new_cases))

for (i in 1:nrow(new_cases)){
  if (X[i] %in% to_plot){
          new_labs[i] <- X[i]
          }
  else{ 
          new_labs[i]<-"."
          }
}

```

Plotting with minimal selection of countries

```{r}
plot(som_model,
     type = "mapping",
     labels = new_labs)
```
```{r}
plot(som_model, type="dist.neighbours")
```

Also referencing: 

https://www.shanelynn.ie/self-organising-maps-for-customer-segmentation-using-r/



Now clustering on the SOM model?


```{r}
mydata <- som_model$codes 
wss <- (nrow(new_cases)-1)*sum(apply(new_cases,2,var)) 
for (i in 2:15) {
  wss[i] <- sum(kmeans(new_cases, centers=i)$withinss)
}
plot(wss)
```

And here: https://rpubs.com/AlgoritmaAcademy/som


```{r}
# Visualising cluster results
## use hierarchical clustering to cluster the codebook vectors
som_cluster <- cutree(hclust(dist(som_model$codes[[1]])), 6)


# plot these results:

plot(som_model, type="mapping",  main = "Clusters",
      bgcol = rainbow(6)[som_cluster]) 
add.cluster.boundaries(som_model, som_cluster)
```


```{r}
country_clusters <- data.frame(
        rownames(new_cases), som_model$unit.classif
              )
```


```{r}
ab <- data.frame(som_cluster)
ab$node_name = sapply(rownames(ab), substring, 2,)
```


```{r}

c6_cases_norm <- merge(country_clusters, ab, by.x="som_model.unit.classif", by.y="node_name")

colnames(c6_cases_norm) <- c("som_node", "country_name", "som_cluster")

c6_cases_norm$country_name <- gsub("\\.", " ",
  c6_cases_norm$country_name)

c6_cases_norm

```

```{r}
write.csv(c6_cases_norm,  "data/c6_cases_norm.csv")
```



```{r}
c6_cases_norm %>% arrange(som_cluster)
```


```{r}

df2 <- data.frame(apply(
        df %>% dplyr::select(-Turkmenistan, -Kiribati, -Tonga)
        , 2, normalize))

df2$date <- rownames(df2)

library(reshape2)
df2 <- melt(df2, id=c("date"))
colnames(df2) <- c("date", "country_name", "new_confirmed")
df2$date = as.Date(df2$date)
```


```{r}
country_clusters
```



```{r}
CLUST_NUM = 15

country_list <- country_clusters[country_clusters$som_model.unit.classif == CLUST_NUM, "rownames.new_cases."]


g = ggplot() +
  geom_line(data = df2[df2$country_name %in% country_list,], 
            aes(x = date, y = new_confirmed, 
                color = country_name)) +
  theme_bw() + 
        ylab("New confirmed") + 
        labs(title = "COVID new cases (scaled)")

plotly::ggplotly(g)
```





TODO: still need to map the nodes to clusters

------



Matrixfy from scratch, bc we need that

```{r}
data <- read.csv("data/data_smoothed_standardised.csv")
data$date <- as.Date(data$date)
data

```

```{r}

library(reshape2)
VALUE_COL = "new_confirmed_norm"

data2 <- data[c("date", "country_name", VALUE_COL)]

country_matrix <- dcast(data2, country_name ~ date, value.var=VALUE_COL)
country_matrix[is.na(country_matrix)] <- 0

rownames(country_matrix) <- country_matrix$country_name
country_matrix <- subset(country_matrix, select=-country_name)

som_model <- som(
        as.matrix(country_matrix), 
        grid=somgrid(xdim = 4, ydim=4, 
                     topo="hexagonal"), 
        rlen=100, 
        alpha=c(0.05,0.01), 
        keep.data = TRUE,)
```


```{r}
plot(som_model,
     type = "mapping",
     # bgcol = "lightgray",
     labels = rownames(new_cases))
```


```{r}
som_cluster <- cutree(hclust(dist(som_model$codes[[1]])), 6)


# plot these results:

plot(som_model, type="mapping",  main = "Clusters",
      bgcol = rainbow(6)[som_cluster]) 
add.cluster.boundaries(som_model, som_cluster)
```

```{r}
country_clusters <- data.frame(
        rownames(country_matrix), som_model$unit.classif
              )
```


```{r}
ab <- data.frame(som_cluster)
ab$node_name = sapply(rownames(ab), substring, 2,)
```


```{r}

c6_cases_norm <- merge(country_clusters, ab, by.x="som_model.unit.classif", by.y="node_name")

colnames(c6_cases_norm) <- c("som_node", "country_name", "som_cluster")

c6_cases_norm$country_name <- gsub("\\.", " ",
  c6_cases_norm$country_name)

c6_cases_norm

```



```{r}
CLUST_NUM = 6

country_list <- c6_cases_norm[c6_cases_norm$som_cluster == CLUST_NUM, "country_name"]


g = ggplot() +
  geom_line(data = data2[data2$country_name %in% country_list,], 
            aes_string(x = "date", y = VALUE_COL, 
                color = "country_name")) +
  theme_bw() + 
        ylab(VALUE_COL) + 
        labs(title = "COVID new cases (scaled)")

plotly::ggplotly(g)
```





```{r}
write.csv(c6_cases_norm,  "data/c6_cases_norm.csv")
```






---------------

## SOM on Stringency 

Now we do an SOM on stringency


```{r}

VALUE_COL = "stringency_index"

data3 <- data[c("date", "country_name", VALUE_COL)]
```





```{r}
country_matrix <- dcast(data3, country_name ~ date, value.var=VALUE_COL)
country_matrix[is.na(country_matrix)] <- 0
```



```{r}
rownames(country_matrix) <- country_matrix$country_name
country_matrix <- subset(country_matrix, select=-country_name)
```





```{r}
som_model <- som(
        as.matrix(country_matrix), 
        grid=somgrid(xdim = 4, ydim=4, 
                     topo="hexagonal"), 
        rlen=100, 
        alpha=c(0.05,0.01), 
        keep.data = TRUE,)
```


```{r}
plot(som_model,
     type = "mapping",
     # bgcol = "lightgray",
     labels = rownames(new_cases))
```


```{r}
som_cluster <- cutree(hclust(dist(som_model$codes[[1]])), 6)


# plot these results:

plot(som_model, type="mapping",  main = "Clusters",
      bgcol = rainbow(6)[som_cluster]) 
add.cluster.boundaries(som_model, som_cluster)
```


```{r}
country_clusters <- data.frame(
        rownames(country_matrix), som_model$unit.classif
              )
ab <- data.frame(som_cluster)
ab$node_name = sapply(rownames(ab), substring, 2,)
```


```{r}
ab
```


```{r}

c6_stringency_norm <- merge(country_clusters, ab, by.x="som_model.unit.classif", by.y="node_name")

colnames(c6_stringency_norm) <- c("som_node", "country_name", "som_cluster")

c6_stringency_norm$country_name <- gsub("\\.", " ",
  c6_stringency_norm$country_name)
c6_stringency_norm$value_col = VALUE_COL
c6_stringency_norm

```



```{r}
CLUST_NUM = 3

country_list <- c6_stringency_norm[c6_stringency_norm$som_cluster == CLUST_NUM, "country_name"]


g = ggplot() +
  geom_line(data = data3[data3$country_name %in% country_list,], 
            aes_string(x = "date", y = VALUE_COL, 
                color = "country_name")) +
  theme_bw() + 
        ylab(VALUE_COL) + 
        labs(title = "COVID Govt Stringency")

plotly::ggplotly(g)
```


```{r}
write.csv(c6_stringency_norm, "data/c6_stringency.csv", row.names=FALSE)
```


------




## World Map Cluster Time

We've generated these clusters! Now it's time to visualise it on the world map.



```{r}
world_map <- map_data("world2")

ggplot(world_map, aes(x = long, y = lat, group = group)) + 
  geom_polygon(fill = "white", colour = "gray") + 
  theme(legend.position = "bottom", aspect.ratio = 0.6) + 
  ggtitle("Map of World")
```


-------


## Designing a New Cluster Dataset for different iterations of cluster 6

Different types we need for new cases:
- Base
- Standardised
- Smoothed
- Standardised and smoothed


And repeat for stringency.



```{r}

library(reshape2)
```


```{r}
## Define Constants
set.seed(123)

VALUE_COL = "new_confirmed_smooth"  # cases column to use



```

```{r}
data <- read.csv("data/data_smoothed_standardised.csv")
data$date <- as.Date(data$date)

data2 <- data[c("date", "country_name", VALUE_COL)]
country_matrix <- dcast(data2, country_name ~ date, value.var=VALUE_COL)
country_matrix[is.na(country_matrix)] <- 0

rownames(country_matrix) <- country_matrix$country_name
country_matrix <- subset(country_matrix, select=-country_name)

som_model <- som(
        as.matrix(country_matrix), 
        grid=somgrid(xdim = 4, ydim=4, 
                     topo="hexagonal"), 
        rlen=100, 
        alpha=c(0.05,0.01), 
        keep.data = TRUE,)


plot(som_model,
     type = "mapping",
     labels = rownames(new_cases))


som_cluster <- cutree(hclust(dist(som_model$codes[[1]])), 6)


# plot these results:

plot(som_model, type="mapping",  main = "Clusters",
      bgcol = rainbow(6)[som_cluster]) 
add.cluster.boundaries(som_model, som_cluster)


# Map to og 6 clusters
country_clusters <- data.frame(
        rownames(country_matrix), som_model$unit.classif
              )
ab <- data.frame(som_cluster)
ab$node_name = sapply(rownames(ab), substring, 2,)  # Nodes are given by V1, V2, etc, so take substring

ab
country_clusters


c6_cases_norm <- merge(country_clusters, ab, by.x="som_model.unit.classif", by.y="node_name")
colnames(c6_cases_norm) <- c("som_node", "country_name", "som_cluster")
c6_cases_norm$country_name <- gsub("\\.", " ",
  c6_cases_norm$country_name)

c6_cases_norm$value_col = VALUE_COL

c6_cases_norm

```

```{r}
ls <- list(
  new_confirmed=som_model
)
```


```{r}
ls2 <- list
```


```{r}
saveRDS(som_model, file="som_model_test.rds")
```

```{r}
youthought <- readRDS("som_model_test.rds")
```


```{r}
youthought


ay <- cutree(hclust(dist(youthought$codes[[1]])), 6)


# plot these results:

plot(youthought, type="mapping",  main = "Clusters",
      bgcol = rainbow(6)[ay]) 
add.cluster.boundaries(youthought, ay)
```



```{r}
unique(country_clusters$som_model.unit.classif) 
```


```{r}
write.csv(c6_cases_norm, paste("data/c6_cases_", VALUE_COL, ".csv", sep=""),
          row.names=FALSE)
```




```{r}
CLUST_NUM = 3

country_list <- c6_cases_norm[c6_cases_norm$som_cluster == CLUST_NUM, "country_name"]


g = ggplot() +
  geom_line(data = data2[data2$country_name %in% country_list,], 
            aes_string(x = "date", y = VALUE_COL, 
                color = "country_name")) +
  theme_bw() + 
        ylab(VALUE_COL) + 
        labs(title = "COVID new cases (scaled)")

plotly::ggplotly(g)
```

## Combine Clusters DFs


```{r}
read.csv("data/c6_cases_ new_confirmed .csv")
```


```{r}


bound <-read.csv(paste("data/", "c6_cases_new_confirmed.csv", sep=""))
for (i in c(
  # "c6_cases_new_confirmed.csv",
  "c6_cases_new_confirmed_pop.csv",
  "c6_cases_new_confirmed_pop_smooth.csv",
  "c6_cases_new_confirmed_smooth.csv",
  "c6_stringency.csv")) {
  
  results <- read.csv(paste("data/", i, sep=""))
  print(results)
  bound <- rbind(bound, results)
}
```

```{r}
bound
```


```{r}
write.csv(bound, "data/c6_country_clusters_cache.csv",
          row.names=FALSE)
```


```{r}

ggplot(
  
)

```



## Saving SOM models to thing



```{r}
## Define Constants
set.seed(123)

VALUE_COL = "new_confirmed"  # cases column to use



```

```{r}

generate_som <- function(VALUE_COL) {
  
  data <- read.csv("data/data_smoothed_standardised.csv")
  data$date <- as.Date(data$date)
  
  data2 <- data[c("date", "country_name", VALUE_COL)]
  country_matrix <- dcast(data2, country_name ~ date, value.var=VALUE_COL)
  country_matrix[is.na(country_matrix)] <- 0
  
  rownames(country_matrix) <- country_matrix$country_name
  country_matrix <- subset(country_matrix, select=-country_name)
  

  som_model <- som(
          as.matrix(country_matrix), 
          grid=somgrid(xdim = 4, ydim=4, 
                       topo="hexagonal"), 
          rlen=100, 
          alpha=c(0.05,0.01), 
          keep.data = TRUE,)
  
  som_cluster <- cutree(hclust(dist(som_model$codes[[1]])), 6)
  country_clusters <- data.frame(
        rownames(country_matrix), som_model$unit.classif
              )
  ab <- data.frame(som_cluster)
  ab$node_name = sapply(rownames(ab), substring, 2,) 
  c6_cases_norm <- merge(country_clusters, ab, by.x="som_model.unit.classif", by.y="node_name")
  colnames(c6_cases_norm) <- c("som_node", "country_name", "som_cluster")
  c6_cases_norm$country_name <- gsub("\\.", " ",
    c6_cases_norm$country_name)
  
  print(c6_cases_norm)
  
  plot(som_model, type="mapping",  main = "Clusters",
    bgcol = rainbow(6)[som_cluster],
    cex=0.6
    , labels=rownames(country_matrix)
    )
  add.cluster.boundaries(som_model, som_cluster)
  

  return(som_model)
  
}


```

```{r}

for (i in c(
  "new_confirmed",
  "new_confirmed_pop",
  "new_confirmed_pop_smooth",
  "new_confirmed_smooth",
  "stringency_index")) {
  som_mod <- generate_som(i)
  
  saveRDS(som_mod, paste("data/c6_sommodel_", i, ".rds", sep=""))
  }

```



```{r}
som_model_new_confirmed <-  generate_som("new_confirmed")
```

```{r}

```



```{r}
saveRDS(som_model_new_confirmed, "data/c6_sommodel_new_confirmed_AHHHHH.rds")
```


```{r}
som_model_new_confirmed

som_cluster <- cutree(hclust(dist(som_model_new_confirmed$codes[[1]])), 6)

plot(som_model_new_confirmed, type="mapping",  main = "Clusters",
    bgcol = rainbow(6)[som_cluster], labels=sort(rownames(country_matrix)),
    cex=0.8) 
add.cluster.boundaries(som_model_new_confirmed, som_cluster)
```
```{r}
hmm <- readRDS("data/c6_sommodel_new_confirmed.rds")
som_cluster <- cutree(hclust(dist(hmm$codes[[1]])), 6)
plot(hmm, type="mapping",  main = "Clusters",
  bgcol = rainbow(6)[som_cluster],
  cex=0.6
  , labels=sort(rownames(country_matrix))
  )
add.cluster.boundaries(hmm, som_cluster)

```






## Quick merge between cluster cache and key


```{r}
hmm = read.csv("data/c6_country_clusters_cache.csv")

data <- read.csv("data/data_smoothed_standardised.csv")

```



```{r}
data
```

```{r}
country_keyname = unique(data[,c("country_name", "key")])
```



```{r}
write.csv(merge(hmm, country_keyname, on="country_name"),
          "data/c6_country_clusters_cache2.csv",
          row.names = FALSE
          )
```


```{r}
model <- readRDS("data/c6_sommodel_new_confirmed.rds")
model
```

```{r}
        # Visualising cluster results
        som_cluster <- cutree(hclust(dist(model$codes[[1]])), 6)

        plot(model, type="mapping",  main = "Clusters",
            bgcol = rainbow(6)[som_cluster]) 
        add.cluster.boundaries(model, som_cluster)
```



```{r}
cov_data <- read.csv("data/data_smoothed_standardised.csv")
saveRDS(cov_data, "data/cov_data.rds")
```


----


# Time to check output

```{r}
new_confirmed <- readRDS("data/c6_RERUN_new_confirmed.rds")
new_confirmed
```

```{r}


clust <- cutree(hclust(dist(new_confirmed$model$codes[[1]])), 6)

plot(new_confirmed$model, type="mapping",  main = "Clusters",
     bgcol = rainbow(6)[clust],
     cex=0.6
     , labels=sort(unique(new_confirmed$mappings$country_name))
)
add.cluster.boundaries(new_confirmed$model, clust)
```


```{r}
ls <- readRDS("data/c6_RERUN_stringency_index.rds")
ls
```

